---
title: "wk_06_widgets_AS"
author: "Alex Smith"
date: "February 13, 2016"
output: html_document
---

### `googleVis`: ..., geo, pie, tree, motion, ...

The `googleVis` package ports most of the [Google charts](https://developers.google.com/chart/interactive/docs/gallery) functionality.

For every R chunk must set option `results='asis'`, and once before any googleVis plots, set `op <- options(gvis.plot.tag='chart')`.

- [Rmarkdown and googleVis](https://cran.r-project.org/web/packages/googleVis/vignettes/Using_googleVis_with_knitr.html)

- [googleVis examples](https://cran.r-project.org/web/packages/googleVis/vignettes/googleVis_examples.html)

#### `gvisLineChart`

Please note that the Motion Chart is only displayed when hosted on a
web server, or if placed in a directory which has been added to the 
trusted sources in the [security settings of Macromedia]
(http://www.macromedia.com/support/documentation/en/flashplayer/help/settings_manager04.html). 
See the googleVis package vignette for more details. 

```{r gvisLineChart, results='asis'}
suppressPackageStartupMessages({
  library(googleVis)  # install.packages('googleVis')
})
#### `gvisMotionChart`

```{r MotionChartExample, results='asis', tidy=FALSE}
M <- gvisMotionChart(Fruits, 'Fruit', 'Year',
         options=list(width=400, height=350))
plot(M)
```

