---
title: "wk_06_widgets_AS"
author: "Alex Smith"
date: "February 13, 2016"
output: html_document
---
#### `gvisLineChart`

```{r gvisLineChart, results='asis'}
suppressPackageStartupMessages({
  library(googleVis)  # install.packages('googleVis')
})

# must set this option for googleVis charts to show up
op <- options(gvis.plot.tag='chart')

df=data.frame(
  country = c("US", "GB", "BR"), 
  val1    = c(10, 13, 14), 
  val2    = c(23, 12, 32))
Line <- gvisLineChart(df)
plot(Line)
```


line chart examples:

- [What's Really Warming the World? Climate deniers blame natural factors; NASA data proves otherwise](http://www.bloomberg.com/graphics/2015-whats-warming-the-world/)
- [**A Year in Search 2015,, - Google Trends](https://www.google.com/trends/story/2015_US): subtle sparklines


#### `gvisTreeMap`

```{r gvisTreeMap, results='asis'}
Tree <- gvisTreeMap(Regions,  
                    "Region", "Parent", 
                    "Val", "Fac", 
                    options=list(fontSize=16))
plot(Tree)
```

- treemap example: [The President's 2016 Budget Proposal | Visual.ly](http://visual.ly/presidents-2016-budget-proposal?view=true)

#### `gvisMotionChart`

Please note that the Motion Chart is only displayed when hosted on a
web server, or if placed in a directory which has been added to the 
trusted sources in the [security settings of Macromedia]
(http://www.macromedia.com/support/documentation/en/flashplayer/help/settings_manager04.html). 
See the googleVis package vignette for more details. 

```{r MotionChartExample, results='asis', tidy=FALSE}
M <- gvisMotionChart(Fruits, 'Fruit', 'Year',
         options=list(width=400, height=350))
plot(M)
```
