---
title: "Exploring Costa Rican Red Snapper landings through time"
author: "Fisheries Management Group - Caio Faro, Alex Smith, Juliana Herrera"
date: "March 11, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction: Fisheries Management

###**Interactive Examples**

Sea Around Us  <http://www.seaaroundus.org/data/#/spatial-catch>.

## Introduction: Costa Rica
Caio's map without bubles

##Introduction: Fisheries in Costa Rica
Caios map with bubbles, talk about the landing sites

## Objective

- Analize the changes in landings of Red Snapper through time
- Explore the relationship between prices and landings through time 


## Data Sources

- INCOPESCA
  - Landings
  - Prices

```{r cars, echo = TRUE}
summary(cars)
```

## Methods

add function to filter data set (ALEX)
```{r pressure}
plot(pressure)
```

## Results
Landings and prices (interactive plots JULI)

## Results
Regional chart data motion -bubble graph ALEX

## Challenges
that thing with plyr

bubble plot html problem (ALEX)

## Shiny!