---
title: "Exploring Costa Rican Red Snapper landings through time"
author: "Fisheries Management Group - Caio Faro, Alex Smith, Juliana Herrera"
date: "March 11, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction: Fisheries Management

###**Interactive Examples**

[Sea Around Us](http://www.seaaroundus.org/data/#/spatial-catch).

## Introduction: Costa Rica
Caio's map without bubles

##Introduction: Fisheries in Costa Rica
Caios map with bubbles, talk about the landing sites

## Objective

- Analize the changes in landings of Red Snapper through time
- Explore the relationship between prices and landings through time 


## Data Sources

- INCOPESCA
  - Landings
  - Prices

```{r cars, echo = TRUE}
summary(cars)
```

## Methods

add function to filter data set (ALEX)
```{r pressure}
plot(pressure)
```

## Results
Landings and prices (interactive plots JULI)
```{r load, echo=FALSE, include =FALSE}
#==PART 2===========================================================
# Loading data from excel and correction for inflation based on the Consumer Price Index (CPI)
#===================================================================

data <- read_excel('./data/prices.xlsx') 
colnames(data) <- tolower(colnames(data))
CPI <- read_excel('./data/CPI2.xlsx') # Consumer Price Index - source: World Bank
# CPIm <- read_excel('CPI.xlsx') # Consumer Price Index by month : World Bank
catch <- read_excel('./data/For CRSeafood code/catches.xlsx') %>%
  gather('month','catch',Jan:Dec) %>%
  select(-catch, -month) %>% 
  filter(Region != 'Mar Caribe', Species == c('Pargo', 'Pargo Seda')) %>% 
  group_by(Region, Year) %>% 
  summarise(sum=sum(Total)) %>% # total catch of pargo and pargo seda per region and year
  select(region=Region, year=Year, sum)

yrprices <- (data) %>%
  select(-fishery, -spp) %>%
  group_by(year) %>% 
  summarize(meanprice = mean(price, na.rm=T)) %>% 
  left_join(CPI, by = 'year') %>%
  mutate(corrected = meanprice/cpi*cpi[24])
head(yrprices) # corrected prices for inflation. Baseline year = 2013

write.csv(yrprices, file = "prices.csv")

# p <-ggplot(yrprices) + ## Graph of corrected prices per year
#   geom_line(aes(x=year, y=meanprice, colour = 'meanprice')) +
#   geom_line(aes(x=year, y=corrected, colour ='corrected')) +
#   theme_bw()
# p


```

```{r spread, echo=F}
## Using dygraphs
Tprices <-  yrprices %>% 
  select(-cpi)
colnames(Tprices) <- toupper(colnames(Tprices))

# Interactive graph for prices only
dygraph(Tprices) %>% 
   dyRangeSelector() %>% 
  dyAxis("y", label = "Price (colones/kg)") %>% 
  dySeries('CORRECTED', strokeWidth = 3) %>% 
  dySeries('MEANPRICE', strokeWidth = 3)
```

## Results
Regional chart data motion -bubble graph ALEX

## Challenges
that thing with plyr

bubble plot html problem (ALEX)

## Shiny!

## General problems we ALL face
- It's a parenthesis... probably
- You're missing a comma!
- Should that be capitalized?
- Check your working directory
- did you load the library?
- Is your object the right class?
- Are you using plyr? it's EVIL KANIVAL