---
title: "Exploring Costa Rican Snapper Landings Through Time"
author: "Fisheries Management Group - Caio Faro, Alex Smith, Juliana Herrera"
date: "March 11, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readr)
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyr)
library(dygraphs)
library(leaflet)
library(ggvis)
library(DT)

suppressPackageStartupMessages({
  library(googleVis)  # install.packages('googleVis')
})

```

<<<<<<< HEAD
## Introduction: Fisheries Management

<<<<<<< HEAD

=======
Engaging the public
>>>>>>> d4a2fee383577a877151342aa5b8fa9e12371123

###**Interactive Examples**

[Sea Around Us](http://www.seaaroundus.org/data/#/spatial-catch)

[Ocean Health Index](http://www.oceanhealthindex.org/region-scores)

The Pew Charitable Trusts - [Importance of Predators](http://www.pewtrusts.org/en/multimedia/video/2013/fisheries-management-its-better-with-predators)

## Introduction: Costa Rica
Caio's map without bubbles

##Introduction: Fisheries in Costa Rica
=======
## Fisheries in Costa Rica
>>>>>>> c5edd92bed93dae134e8fad047a8b892774a7db6
Caios map with bubbles, talk about the landing sites

## Objectives

- Analize the changes in landings of Snapper through time
- Explore the relationship between prices and landings through time 

![]()

## Data Sources

- INCOPESCA
  - Landings
  - Prices
  
## Our Data
Annual Catch Data for ***Pargo*** and ***Pargo Seda*** in 4 Pacfic Regions (1990-2013)
```{r html data table ALEX}

catchT=read.csv('../data/For CRSeafood code/motioncatch.csv')
filt=list(position = 'top', clear = TRUE, plain = FALSE)

datatable(catchT, options = list(pageLength = 5, dom = 'tip'), filter =filt, class = "compact")
```

<<<<<<< HEAD
Data source: INCOPESCA

## Methods
=======
## Results: Landings through time

```{r, echo=FALSE, warning=FALSE}

catch <- read_excel('../data/For CRSeafood code/catches.xlsx') %>%
  gather('month','catch',Jan:Dec) %>%
  select(-catch, -month) %>% 
  filter(Region != 'Mar Caribe', Species == c('Pargo', 'Pargo Seda')) %>% 
  group_by(Region, Year) %>% 
  summarise(sum=sum(Total)) %>% # total catch of pargo and pargo seda per region and year
  select(region=Region, year=Year, sum)

landings <- as.data.frame(catch)

 ggplot(landings, aes(x = year, y = sum, col = region)) +
  geom_line(size=1.5) +
  ylab("Total Landings (kg)")+
  xlab("Year")+
  theme_bw()

```
>>>>>>> 74a6bce5b38f9b48cb950ce9916e01979ee32a0c

## Results: Price through time

```{r load, echo=FALSE, include =FALSE}
#==PART 2===========================================================
# Loading data from excel and correction for inflation based on the Consumer Price Index (CPI)
#===================================================================

data <- read_excel('../data/prices.xlsx') 
colnames(data) <- tolower(colnames(data))
CPI <- read_excel('../data/CPI2.xlsx') # Consumer Price Index - source: World Bank
# CPIm <- read_excel('CPI.xlsx') # Consumer Price Index by month : World Bank

yrprices <- (data) %>%
  select(-fishery, -spp) %>%
  group_by(year) %>% 
  summarize(meanprice = mean(price, na.rm=T)) %>% 
  left_join(CPI, by = 'year') %>%
  mutate(corrected = meanprice/cpi*cpi[24])
head(yrprices) # corrected prices for inflation. Baseline year = 2013

write.csv(yrprices, file = "prices.csv")

# p <-ggplot(yrprices) + ## Graph of corrected prices per year
#   geom_line(aes(x=year, y=meanprice, colour = 'meanprice')) +
#   geom_line(aes(x=year, y=corrected, colour ='corrected')) +
#   theme_bw()
# p


```

```{r spread, echo=F}
## Using dygraphs
Tprices <-  yrprices %>% 
  select(-cpi)
colnames(Tprices) <- toupper(colnames(Tprices))

# Interactive graph for prices only
dygraph(Tprices) %>% 
   dyRangeSelector() %>% 
  dyAxis("y", label = "Price (colones/kg)") %>% 
  dySeries('CORRECTED', strokeWidth = 3) %>% 
  dySeries('MEANPRICE', strokeWidth = 3)
```

## Results
Regional chart data motion -bubble graph ALEX
```{r bubble chart, echo=FALSE}
op <- options(gvis.plot.tag='chart')  

motioncatch=group_by(catchT, Region,Year) %>%
              summarise(
                
                Annual_Catch=sum(Catch)
              ) 

InitialState='{"playDuration":15000,"xLambda":1,"yAxisOption":"2","iconType":"BUBBLE","xAxisOption":"_TIME","xZoomedDataMax":1356998400000,"yZoomedDataMin":506,"orderedByX":false,"yZoomedIn":false,"sizeOption":"2","yLambda":0,"uniColorForNonSelected":false,"showTrails":true,"iconKeySettings":[],"xZoomedDataMin":631152000000,"orderedByY":false,"nonSelectedAlpha":0.4,"duration":{"multiplier":1,"timeUnit":"Y"},"colorOption":"_UNIQUE_COLOR","dimensions":{"iconDimensions":["dim0"]},"time":"1990","xZoomedIn":false,"yZoomedDataMax":634463};'

C= gvisMotionChart(motioncatch, idvar='Region', timevar = 'Year',  sizevar = 'Annual_Catch', colorvar = 'Region',  options=list(state=InitialState))


plot(C)


```

##Informatics and Fisheries Management

Engaging the public

###**Interactive Examples**

[Sea Around Us](http://www.seaaroundus.org/data/#/spatial-catch).

The Pew Charitable Trusts - [Importance of Predators](http://www.pewtrusts.org/en/multimedia/video/2013/fisheries-management-its-better-with-predators)

## Challenges
that thing with plyr

bubble plot html problem (ALEX)
```{r bubble problem ALEX, eval=FALSE, echo=TRUE, collapse=T, tidy=T}


InitialState='{"playDuration":15000,"xLambda":1,"yAxisOption":"2","iconType":"BUBBLE","xAxisOption":"_TIME","xZoomedDataMax":1356998400000,"yZoomedDataMin":506,"orderedByX":false,"yZoomedIn":false,"sizeOption":"2","yLambda":0,"uniColorForNonSelected":false,"showTrails":true,"iconKeySettings":[],"xZoomedDataMin":631152000000,"orderedByY":false,"nonSelectedAlpha":0.4,"duration":{"multiplier":1,"timeUnit":"Y"},"colorOption":"_UNIQUE_COLOR","dimensions":{"iconDimensions":["dim0"]},"time":"1990","xZoomedIn":false,"yZoomedDataMax":634463};'

C= gvisMotionChart(motioncatch, idvar='Region', timevar = 'Year',  sizevar = 'Annual_Catch', colorvar = 'Region',  options=list(state=InitialState))


plot(C)


```

## Shiny!

## General problems we ALL face
- It's a parenthesis... probably
- You're missing a comma!
- Should that be capitalized?
- Check your working directory
- Did you load the library?
- Is your object the right class?
- Are you using plyr? it's EVIL KANIVAL

!(https://c1.staticflickr.com/5/4007/5143180652_3f23fe8d95.jpg)
